\documentclass[journal]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts}
\usepackage{array}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{stfloats}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage[hidelinks]{hyperref}

\usetikzlibrary{arrows.meta,positioning,fit}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcommand{\cmark}{Yes}
\newcommand{\xmark}{No}
\newcommand{\partialmark}{Partial}

\hyphenation{Vision-Language-Action multi-modal robot-ics IEEE-Xplore}

\begin{document}

\title{Attacks on Vision-Language-Action Robotic Systems: A Systematic Evidence Review}

\input{author}

\markboth{Anonymous \MakeLowercase{\textit{et al.}}: Attacks on Vision-Language-Action Models}{}

\maketitle

\begin{abstract}
Vision-Language-Action (VLA) models turn perception and language into executable robot behavior. This creates an attack surface that is not reducible to adversarial image classification, text-only jailbreaking, conventional robot robustness, or broad VLA safety: a small visual, linguistic, training-time, or system-level intervention can propagate through grounding, planning, action representation, controller execution, and recovery. This survey reviews how VLA attacks emerge in closed-loop robotic systems, how the literature has evolved from adversarial vision, prompt injection, and robot-policy attacks toward direct VLA patches, black-box instruction attacks, action-level backdoors, trajectory manipulation, reasoning-chain hijacking, and privacy attacks, and what current evidence can and cannot support. We use a structured evidence review to separate direct VLA attacks from embodied-adjacent evidence and broader contextual literature, while keeping the main narrative focused on attack mechanisms and robotic consequences. Because attack success is not yet used consistently in VLA safety work, we formalize typed VLA attack success as a rollout-level violation predicate tied to attacker assumptions, benign baselines, validation tier, and consequence. We introduce an orthogonal attack map spanning lifecycle, entry surface, target component, target representation, security property, knowledge level, access channel, privilege level, validation tier, and consequence. Across studies, the evidence shows rapid growth but also clear limits: direct VLA results remain concentrated around OpenVLA/LIBERO-style settings, physical validation is uneven, and controller, recovery, and pre/post-controller traces are rarely reported. We close with a robotics-oriented evaluation protocol and a research roadmap for physically grounded, action-representation-aware, and recovery-aware VLA attack studies.
\end{abstract}

\begin{IEEEkeywords}
Vision-Language-Action Attacks, Robot Foundation Models, Systematic Evidence Review, Embodied Security, Physical Attack Evaluation.
\end{IEEEkeywords}

\input{sections/1_Introduction}
\input{sections/2_BackgroundThreatModel}
\input{sections/3_Taxonomy}
\input{sections/4_PerceptionGrounding}
\input{sections/5_InstructionsPlanning}
\input{sections/6_ActionExecution}
\input{sections/7_MemoryToolEnvironment}
\input{sections/12_Conclusion}

{
\scriptsize
\bibliographystyle{ieeetr}
\bibliography{main}
}

\vfill

\end{document}
