\section{Conclusion}
\label{sec:conclusion}

VLA attacks should be judged at the level of closed-loop consequences, not isolated model outputs. An adversarial effect may enter through perception, language, memory, tools, environment state, action representations, or human interaction. It may first appear as a changed object binding, plan, reasoning trace, action token, chunk, trajectory, tool call, or memory entry. The security claim becomes a robotics claim only when that change crosses the controller boundary, interacts with feedback and recovery, and reaches a task, safety, authorization, privacy, or persistence violation.

The emerging literature already shows that direct VLA systems can be attacked through patches, black-box instructions, action freezing, action-level and long-horizon backdoors, transferable and 3D object perturbations, reasoning-chain hijacking, trajectory redirection, and membership inference. Embodied-adjacent work usefully motivates risks around planners, tools, sensors, middleware, and HRI, while contextual work explains mechanisms and evaluation design. These levels should remain separate: task failure, targeted action, safety violation, privacy leakage, unauthorized skill use, persistence, and recovery failure are different predicates and should not be collapsed into one ASR.

For VLA systems, an attack claim is ultimately a claim about what survives the robot loop. Future work should test whether attacks transfer across action representations, remain consequential under physical or HIL recovery protocols, and require privacy-aware metrics rather than task ASR. It should also evaluate adaptive attackers against deployed guardrails while releasing artifacts in forms that support defense without enabling misuse.
