\section{Evolution of VLA Attack Mechanisms}
\label{sec:evidence-synthesis}

The field has developed in waves. The first grew out of adversarial vision and physical patches: if camera inputs can be perturbed, robot perception and grounding may change. The second drew on prompt injection, embodied jailbreaks, and robot-policy attacks, where language and task context steer plans, skills, and sequential decisions. The current wave attacks VLA systems directly: patches are optimized for action deviation, prompts are evaluated by rollout behavior, backdoors target primitives and chunks, reasoning traces are treated as control-relevant state, trajectories are redirected, and action traces become privacy signals. This section follows that evolution by mechanism rather than by modality alone. Table~\ref{tab:attack-literature-summary} summarizes representative VLA attack methods and their evaluation settings.

\input{tables/literature_matrix_table}

\subsection{Training-Time, Supply-Chain, and Backdoor Attacks}

Training-time attacks arise because VLA systems are adapted from large robot datasets, demonstrations, and fine-tuning corpora. A poisoned demonstration set may look normal during training while associating a trigger with a gripper primitive or a small delta-pose drift. At deployment, the robot may still perform clean tasks, yet the trigger causes a release, freeze, wrong-object action, or gradual trajectory shift.

VLA backdoor work has therefore moved beyond generic task disruption. The common mechanism is trigger-conditioned control of a robot-relevant representation while clean behavior remains plausible. Some studies target triggered action deviations or physical-object-conditioned goals \cite{zhou2025badvla,zhou2025goba}. Others bind triggers to reusable primitives, long-horizon sequences, static action states, or chunk-level drift \cite{xu2025dropvla,li2025attackvla,wang2025freezevla,xu2026silentdrift}. The target is no longer a label; it is the command object the robot will execute.

The robotic issue is temporal control. A backdoor that fires a primitive at the wrong decision window differs from one that slowly biases a chunk, and both differ from a trigger that changes a high-level goal. Direct VLA studies are stronger than classifier analogies because they measure action behavior and often clean-task retention, but the settings remain narrow: most results use curated manipulation suites, controlled triggers, and limited embodiments. Current evidence leaves unresolved whether triggers survive data filtering, downstream fine-tuning, tokenized versus continuous action heads, chunked policies, flow/diffusion policies, and real controller clipping. Classic poisoning/backdoor work \cite{biggio2012poisoning,gu2017badnets,liu2018trojaning,carlini2024poisoningweb} and RL backdoors \cite{kiourti2020trojdrl,wang2021backdoorl,cui2024badrl,rathbun2025inception} explain the supply-chain mechanism; direct VLA work is what establishes the robotics claim.

\subsection{Perception, Multimodal, and Physical Attacks}

Visual and physical attacks are the most natural bridge from adversarial ML to VLA robotics, but fooling an image model is not enough. The robotics claim begins when a perturbation survives the perception-to-action chain: a patch or object texture changes which object the robot grounds, shifts the action trajectory, and is then either corrected by feedback or passed to the controller as wrong-object motion \cite{lin2017tactics,sharif2016accessorize,xie2017segdet,sato2021dirtyroad}.

Early VLA patch work made this shift explicit by optimizing action discrepancy, position-aware objectives, and targeted manipulation behavior rather than image classification error \cite{wang2025vlaadversarial}. Other work attacks earlier representations: VLA-Fool studies textual, visual, and cross-modal misalignment, while EDPA and ADVLA target visual embeddings, projected features, attention, and vision--language alignment \cite{yan2025vlafool,xu2025edpa,zhang2025advla}. Transfer-oriented work asks whether shared feature-space, attention, semantic, or visual-proprioception structure can carry attacks beyond one model-specific action head \cite{lu2026patch,fu2026vlahijack}. Physical realism appears in object-bound texture attacks and in partially observable settings where the adversary sees only a short trajectory prefix \cite{chen2026tex3d,wang2026partialpatch}.

The mechanism sequence is now clearer. Action-space patches ask whether visual perturbations can change robot motion. Embedding and attention attacks ask whether the shared multimodal representation is enough when the action head is unknown. Transferable patches ask whether the attack survives changes in architecture, fine-tuning, viewpoint, and sim-to-real setting. 3D texture attacks ask whether the manipulated object can carry the perturbation. Partial-observation attacks ask whether the attack can be planned without knowing the full future rollout.

The unresolved issue is not whether vision can be perturbed; adversarial examples and physical patches have already established that mechanism \cite{szegedy2014intriguing,goodfellow2015explaining,papernot2016limitations,carlini2017towards,brown2017adversarial} \cite{kurakin2017physical,moosavi2017universal,madry2018towards,athalye2018synthesizing,eykholt2018robust}. 3D, detector, LiDAR, and audio-command settings extend the concern beyond flat images \cite{eykholt2018detectors,chen2018shapeshifter,thys2019fooling,xu2020tshirt,xiao2019meshadv} \cite{tu2020lidar,sun2020robustlidar,cao2021invisible,cao2019lidar,carlini2016hiddenvoice} \cite{zhang2017dolphinattack,yuan2018commandersong,sugawara2020lightcommands}. What remains uncertain is how reliably a physical perturbation changes grounding, action, trajectory, and recovery across robot embodiments and controllers.

\subsection{Instruction, Planning, and Reasoning Attacks}

Instruction attacks matter because VLA systems treat language as a control input, not only as a conversation channel. In scene-text hijacking, the user asks for a normal task, but text in the environment or a small edit in the instruction changes object binding, subgoal order, or trajectory preference. A subtler case is command-preserving redirection: the instruction remains close to the intended task while the physical path moves toward a different outcome.

RoboGCG adapts adversarial prompt optimization to VLA control authority and shows that a prompt-level intervention at rollout start can influence action reachability over time \cite{jones2025robogcg}. VLA-Fool includes textual and cross-modal misalignment as well as visual perturbations \cite{yan2025vlafool}. SABER makes the instruction threat model more practical by using a black-box edit process with bounded character, token, or prompt changes and rollout-level feedback across multiple VLA models \cite{wu2026saber}. Q-DIG approaches red-teaming differently, using quality-diversity search to generate natural, task-relevant instructions that expose robot-policy failures and can support robustness improvement \cite{srikanth2026qdig}. Trajectory-level redirection studies show why instruction similarity is not enough: a prompt can remain close to the intended command while redirecting the physical rollout \cite{puthumanaillam2026trajectory}. Prompt and multimodal red-teaming work supplies useful background on optimization, jailbreak transfer, and benchmark design, but robot-level claims require VLA rollouts \cite{perez2022ignore,wei2023jailbroken,chao2023pair,liu2024autodan,mazeika2024harmbench} \cite{shen2024dan,mehrotra2024tap,zhu2023promptbench,li2023deepinception,paulus2025advprompter}.

Reasoning-enabled VLAs add another target between language and action. TRAP shows that CoT or intermediate reasoning traces can be hijacked by physical patches to steer downstream actions without changing the user's instruction \cite{huang2026trap}. Altered-thought studies indicate that entity references in VLA reasoning traces can causally affect robot manipulation success, while non-reasoning control models may be immune to that channel \cite{altered2026thoughts}. CoT-VLA illustrates the constructive side of this interface by using visual chain-of-thought reasoning before action generation \cite{zhao2025cotvla}. ReasonBreak extends the concern to autonomous-driving-style VLA systems, evaluating black-box textual corruptions against reasoning and trajectory safety metrics in closed-loop simulation \cite{teymoorianfard2026reasonbreak}. In these architectures, reasoning text is not merely explanation; it can become state that the controller ultimately acts on.

Embodied jailbreak and planning benchmarks provide useful context. BadRobot, RoboPAIR, and RoboJailBench evaluate embodied jailbreaks where success is measured closer to action than text-only refusal \cite{zhang2025badrobot,robey2025robopair,yeke2026robojailbench}. CHAI, SafeAgentBench, and EARBench add physical command hijacking and risk-aware planning settings \cite{burbano2025chai,yin2024safeagentbench,zhu2024earbench}. Multimodal jailbreak benchmarks further clarify source and modality confusion, but remain contextual unless the outcome reaches robot action \cite{gong2025figstep,niu2024jailbreaking,liu2024mmsafetybench,li2024hades}. Their limitation for this survey is target specificity: when the victim is an LLM-controlled planner rather than a VLA policy, the result motivates the attack surface but does not prove VLA action vulnerability. Direct VLA studies still need to test how instruction and reasoning attacks behave under prompt hierarchy, source labels, physical context changes, and closed-loop recovery.

\subsection{Action-Level and Closed-Loop Attacks}

Action representation is the robotics core of the survey. In an action-freezing attack, the perception and language inputs look plausible, but the policy repeatedly emits a static or blocked action state, causing the robot to stall or fail to release control at the right moment. Chunk drift is subtler: each predicted delta-pose is only slightly biased, but the robot executes a multi-step chunk before new observations can correct the drift.

Autoregressive action-token policies expose token probabilities and discretized gripper or pose bins. A prompt or patch can change which token is selected, and early token errors can propagate through subsequent generation. RT-1, RT-2, OpenVLA, SpatialVLA, UniVLA, and FAST-style tokenizers illustrate tokenized or discretized interfaces whose vocabularies, spatial grids, and detokenizers become attack-relevant objects \cite{brohan2022rt1,brohan2023rt2,kim24openvla,qu2025spatialvla,wang2025unified} \cite{pertsch2025fast,shafiullah2022bet,lee2024vqbet}. Parallel action chunks create a different risk: the model commits to several future actions before new observations can correct the error. ACT, OpenVLA-OFT, RDT-1B, and related high-throughput fine-tuning recipes show why chunk length, temporal ensemble, and parallel decoding are security variables as well as efficiency variables \cite{zhao2023act,kim2025openvlaoft,liu2024rdt,pertsch2025fast}. SilentDrift directly exploits this intra-chunk open-loop property \cite{xu2026silentdrift}.

Waypoint and trajectory policies expose geometric objectives such as end-effector displacement, curvature, unsafe workspace entry, or targeted path redirection \cite{wang2025vlaadversarial,puthumanaillam2026trajectory}. Planner-mediated systems make object binding, generated code, value maps, and waypoint plans inspectable; CLIPort, Transporter Networks, PerAct, Act3D, Code as Policies, ProgPrompt, and VoxPoser show why these intermediate objects can become attack targets \cite{shridhar2022cliport,zeng2020transporter,shridhar2022peract,gervet2023act3d} \cite{liang2023code,singh2023progprompt,huang2023voxposer,goyal2023rvt}. Diffusion and flow policies shift the target from a discrete token to a conditional trajectory distribution, denoising path, velocity field, or sampling process \cite{chi2023diffusion,black2024pi0,ze2024dp3,xian2024diffuseractor,liu2024rdt} \cite{reuss2024mdt,lu2026h3dp}. Generalist policy work defines cross-embodiment assumptions \cite{ghosh2024octo,openx2023,reed2022gato,bousmalis2023robocat,nvidia2025gr00tn1}; robot-data collections expose provenance and embodiment choices that shape transfer claims \cite{khazatsky2024droid,walke2023bridgedata,dasari2019robonet,mandlekar2021robomimic,mandlekar2023mimicgen}.

Whether an action-level attack becomes physically consequential depends on controller and recovery design. A temporal ensemble can damp high-frequency perturbations but may also prolong a poisoned chunk. A high re-planning frequency can recover from one-frame perception attacks yet remain vulnerable to repeated low-amplitude nudges. Controller latency and stale observations can turn a semantically correct command into an unsafe physical action. Safety envelopes and low-level controllers can prevent direct collision while still allowing wrong-object manipulation, unauthorized skill use, privacy leakage, or near-boundary behavior. The difficult case is not only to attack an action representation, but to show that the attack crosses the controller boundary and survives recovery.

\subsection{System, Memory, Tool, HRI, and Persistent Attacks}

VLA deployments will not be isolated policies. They will use tools, maps, object memories, user preferences, middleware, logs, and human instructions. Persistent compromise can arise when a corrupted object memory or user preference is written during one session and retrieved in a later session, causing the robot to choose the wrong object or tool even though the new instruction is benign. Another case is an untrusted tool response that changes a skill argument or authorization boundary.

Tool-agent attacks and indirect prompt injection show how untrusted context can override source boundaries and cause tool misuse or data leakage \cite{greshake2023not,zhan2024injecagent,debenedetti2024agentdojo}. Related benchmarks and attack studies sharpen the same source-confusion problem for software agents \cite{yi2023bipia,ruan2024toolemu}. In robotics, the analogous failure can become an unauthorized skill call, unsafe tool argument, or middleware command. Toolformer, ReAct, Gorilla, and ToolLLM are useful as software-agent substrates, while WebShop and WebArena illustrate interactive task settings \cite{schick2023toolformer,yao2023react,patil2023gorilla,qin2023toolllm,yao2022webshop,zhou2023webarena}. VLA claims still require robot skills or VLA-controlled actions in the loop. ROS/ROS2 and robot cybersecurity work show that middleware, topics, services, node identity, and controller interfaces can be compromised \cite{quigley2009ros,dieber2020penetration,deng2022insecurityros2,neupane2024security,yaacoub2022robotics}. Industrial, surgical, and service-robot studies show analogous weaknesses in deployed platforms \cite{quarta2017industrial,bonaci2015surgical,mayoral2020robot}. HRI work shows that robots can affect human trust, authority, and physical access decisions \cite{booth2017piggybacking,postnikoff2018robot,bainbridge2011benefits,saunderson2021persuasive}.

Memory and privacy broaden the consequence space. RAG and memory poisoning show how retrieved context can be corrupted and reused later \cite{zhong2023poisoning,deng2024pandora,zou2025poisonedrag,cho2024garag}. For VLA robots, multi-session embodied memory compromise remains underdeveloped: poisoned maps, object memories, tool logs, or preferences may survive correction and later redirect behavior. VLA membership-inference work exposes a different risk. Action likelihoods, action errors, temporal motion patterns, attention discrepancies, and trajectory structure can leak training membership \cite{peng2026miavla,luan2026vlaleaks}. This is a direct VLA security problem even when no collision or task failure occurs; evaluation must connect confidentiality, persistence, authorization, and physical task performance.
